---
title: "OB full/skinny R est"
author: "Christine Markwalter"
date: "7/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting set up and reading in data
```{r}
library(tidyverse)
library(lubridate)
library(EpiEstim)


skinny_OB_list <- readRDS("data/generated_tables/OB_dfs/all_skinny_dfs.rds")
full_OB <- read.csv("data/generated_tables/full_OB.csv", row.names = NULL)
malaria_SI <- read.csv("data/generated_tables/malaria_SI.csv") %>%
  filter(t >= 0, p >= 0)
malaria_SI$p[malaria_SI$t == 0] <- 0
```

# Full OB Rt based on pcr-positivity
```{r}
##get incidence for pcr positivity
full_OB_pcrpos <- full_OB %>%
  filter(haplotype_reads > 0) %>%
  group_by(date) %>%
  summarize(I = n())


##include all dates
full_OB_pcrpos$date <- ymd(full_OB_pcrpos$date)
full_OB_dates <- seq(ymd('2017-06-12'),ymd('2018-07-25'),by='days')
full_OB_dates <- data.frame(full_OB_dates) %>%
  rename("date" = full_OB_dates)
full_OB_pcrpos <- left_join(full_OB_dates, full_OB_pcrpos)
full_OB_pcrpos$I[is.na(full_OB_pcrpos$I)] <- 0

##estimate Rt
t_start_full = seq(2, nrow(full_OB_pcrpos)-7)
t_end_full = t_start_full + 7

full_OB_rt <- wallinga_teunis(full_OB_pcrpos$I, 
                        method = 'non_parametric_si', 
                        config = make_config(list(
                         si_distr = malaria_SI$p,
                         t_start = t_start_full,
                         t_end = t_end_full
                       )))

plot(full_OB_rt)

## summarizing Rt results
full_df <- data.frame(t_start = full_OB_rt$R$t_start,
                        est = full_OB_rt$R$`Mean(R)`,
                        lower_ci = full_OB_rt$R$`Quantile.0.025(R)`,
                        upper_ci= full_OB_rt$R$`Quantile.0.975(R)`) %>%
             mutate(date_start = full_OB_pcrpos$date[t_start], date = date_start+3, rcd_limit = "Full OB")
  

```

#Skinny OB Rt calcualtions
```{r}

WT_list <- function(skinny_OB) {
##get incidence for pcr positivity
skinny_OB_pcrpos <- skinny_OB %>%
  filter(haplotype_reads > 0) %>%
  group_by(date) %>%
  summarize(I = n())

##include all dates
skinny_OB_pcrpos$date <- ymd(skinny_OB_pcrpos$date)
skinny_OB_dates <- seq(ymd('2017-06-12'),ymd('2018-07-25'),by='days')
skinny_OB_dates <- data.frame(skinny_OB_dates) %>%
  rename("date" = skinny_OB_dates)
skinny_OB_pcrpos <- left_join(skinny_OB_dates, skinny_OB_pcrpos)
skinny_OB_pcrpos$I[is.na(skinny_OB_pcrpos$I)] <- 0

##estimate Rt
t_start_skinny = seq(2, nrow(skinny_OB_pcrpos)-7)
t_end_skinny = t_start_skinny + 7

skinny_OB_rt <- wallinga_teunis(skinny_OB_pcrpos$I, 
                        method = 'non_parametric_si', 
                        config = make_config(list(
                         si_distr = malaria_SI$p,
                         t_start = t_start_skinny,
                         t_end = t_end_skinny
                       )))

## summarizing Rt results
skinny_df <- data.frame(t_start = skinny_OB_rt$R$t_start,
                        est = skinny_OB_rt$R$`Mean(R)`,
                        lower_ci = skinny_OB_rt$R$`Quantile.0.025(R)`,
                        upper_ci= skinny_OB_rt$R$`Quantile.0.975(R)`) %>%
             mutate(date_start = skinny_OB_pcrpos$date[t_start], date = date_start+3)

return(skinny_df)

}

skinny_Rt_list <- lapply(skinny_OB_list, WT_list)

skinny_Rt_df <- bind_rows(skinny_Rt_list, .id = "rcd_limit")
skinny_Rt_df$rcd_limit <- as.numeric(skinny_Rt_df$rcd_limit)
```




# Plot with all the curves
```{r}
library(RColorBrewer)
myPalette <- colorRampPalette(brewer.pal(3, "Pastel2"))

Rt_plot<- ggplot(NULL, aes(x,y)) +
  geom_line(skinny_Rt_df %>% filter(rcd_limit != 24), mapping = aes(x=date, y=est, color= as.factor(rcd_limit)), size = 0.8) +
  geom_hline(yintercept = 1, linetype="dashed") +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  #scale_color_manual(values = myPalette(24)) +
  #scale_fill_manual(values = myPalette(24)) +
  geom_line(full_df, mapping = aes(x=date, y=est, ymin=lower_ci, ymax=upper_ci), color = "black", size = 1) +
  geom_ribbon(full_df, mapping = aes(x=date, y=est, ymin=lower_ci, ymax=upper_ci), alpha = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "Rt", x = "Date")
  
Rt_plot
```


# Scatter plot comparing Rt estimates for skinny and full OB
```{r}
skinny_Rt_df$rcd_limit <- as.factor(skinny_Rt_df$rcd_limit)
merged <- skinny_Rt_df %>%
  left_join(full_df %>% select(-t_start, -date_start, -rcd_limit) %>% rename("full_est" = "est", "full_lower_ci" = "lower_ci", "full_upper_ci" = "upper_ci"), by = "date")


fullvskinny <- ggplot(merged %>% filter(!is.na(rcd_limit), rcd_limit != 24), aes(x = full_est, y = est, color = rcd_limit)) +
  geom_point() +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  geom_abline(slope=1, intercept=0) +
  scale_x_continuous(limits = c(0,3))+
  scale_y_continuous(limits = c(0,6)) +
  labs(x = "Rt, full dataset", y = "Rt, simulated reactive case detection", color = "Reactive events/month limit") +
  theme_bw()

fullvskinny


p3 <- ggplot(merged %>% filter(rcd_limit != 24) %>% mutate(dist = (est - full_est)) %>% filter(!is.na(dist)), aes(x = date, y = dist, color = rcd_limit)) +
  geom_point() +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  geom_hline(yintercept = 0) +
  labs(x = "Date", y = "\u0394 Rt", color = "rcd limit") +
  theme_bw()

p3

p4<- ggplot(merged %>% filter(!rcd_limit %in% c(1,24)) %>% mutate(dist = (est - full_est)) %>% filter(!is.na(dist)), aes(x = full_est, y = dist/full_est, color = rcd_limit)) +
  geom_point() +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  geom_hline(yintercept = 0) +
  labs(x = "Full dataset Rt", y = "\u0394 Rt/(Full Rt)", color = "rcd limit") +
  theme_bw()

p4

p5<- ggplot(merged %>% filter(!rcd_limit %in% c(1,24)) %>% mutate(dist = (est - full_est)) %>% filter(!is.na(dist)), aes(x = full_est, y = dist, color = rcd_limit)) +
  geom_point() +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  geom_hline(yintercept = 0) +
  labs(x = "Full dataset Rt", y = "\u0394 Rt", color = "rcd limit") +
  theme_bw()
 p5

```

#Correlation between skinny and full
```{r}
merged <- merged %>% 
  filter(!is.na(rcd_limit), rcd_limit != 24)

#do spearman's rho

cor.test(x = merged$full_est, y = merged$est, method = "spearman")


```
#Estimating overall R
```{r}
t_start <- c(2, 61, 350) 
t_end <- c(60, 349, 409)

full_R <- wallinga_teunis(full_OB_pcrpos$I, 
                        method = 'non_parametric_si', 
                        config = make_config(list(
                         si_distr = malaria_SI$p,
                         t_start = t_start,
                         t_end = t_end
                       )))


full_R$R


####For skinny
Overall_WT_list <- function(skinny_OB) {
##get incidence for pcr positivity
skinny_OB_pcrpos <- skinny_OB %>%
  filter(haplotype_reads > 0) %>%
  group_by(date) %>%
  summarize(I = n())

##include all dates
skinny_OB_pcrpos$date <- ymd(skinny_OB_pcrpos$date)
skinny_OB_dates <- seq(ymd('2017-06-12'),ymd('2018-07-25'),by='days')
skinny_OB_dates <- data.frame(skinny_OB_dates) %>%
  rename("date" = skinny_OB_dates)
skinny_OB_pcrpos <- left_join(skinny_OB_dates, skinny_OB_pcrpos)
skinny_OB_pcrpos$I[is.na(skinny_OB_pcrpos$I)] <- 0

##estimate R


skinny_OB_R <- wallinga_teunis(skinny_OB_pcrpos$I, 
                        method = 'non_parametric_si', 
                        config = make_config(list(
                         si_distr = malaria_SI$p,
                         t_start = t_start,
                         t_end = t_end
                       )))

return(skinny_OB_R)

}

skinny_R <- lapply(skinny_OB_list, Overall_WT_list)
temp <- transpose(skinny_R)
skinny_R_df <- bind_rows(temp$R, .id = "rcd_limit")

overall_R_plot <- ggplot(skinny_R_df %>% filter(rcd_limit != 24, t_start == 61), aes(x = `Mean(R)`, xmin = `Quantile.0.025(R)`, xmax = `Quantile.0.975(R)`, y = factor(as.numeric(rcd_limit)))) +
  ggstance::geom_pointrangeh(aes(color = factor(as.numeric(rcd_limit)))) +
  scale_color_viridis_d(begin = 0.3, alpha = 0.8) +
  ggstance::geom_pointrangeh(data = full_R$R %>% filter(t_start == 61), mapping = aes(x = `Mean(R)`, xmin = `Quantile.0.025(R)`, xmax = `Quantile.0.975(R)`, y = factor(24))) +
  scale_y_discrete(limits = factor(c(1:24)), labels = c(1:23, "None")) +
  labs(y = "Montly reactive event limit", x = "Estimated Overall R") +
  theme_bw() +
  theme(legend.position = "none")

overall_R_plot

```


#Manuscript figure
```{r}
OBfig <- ggarrange(overall_R_plot,
          Rt_plot,
          fullvskinny,
          common.legend = TRUE,
          legend = "none",
          nrow = 1,
          labels = c("a","b", "c"))

OBfig

ggsave(plot = OBfig, "manuscript/figures/OBfig.png", height = 4, width = 10)


```





----Stop here---
# Proportion of events captured
```{r}
months_df <- read.csv("data/generated_tables/OB_dfs/skinny_OB-23.csv") %>%
  filter(reactive_event == "yes") %>%
  select(month_year, reactive_id) %>%
  distinct() %>%
  group_by(month_year) %>%
  summarize(total_events = n())


#make columns with the proportion of monthly events captured in that month given the limit i
for(i in 1:23) {
  months_df <- months_df %>% mutate_(.dots = setNames(list(paste0("ifelse(total_events >", i, ",", i, "/total_events,1)")), i))
  }

months_df$month_year <- ym(months_df$month_year)

next_month <- months_df %>%
  mutate(shift_month = month_year - months(1)) %>%
  select(-month_year) %>%
  rename("month_year" = "shift_month", "total_events_nextmonth" = "total_events") %>%
  pivot_longer(2:24, names_to = "rcd_limit", values_to = "prop_events_nextmonth")
  

#Now, I want to get the Rt estimates in there. But first, I will pivot longer
months_df <- months_df %>%
  pivot_longer(3:25, names_to = "rcd_limit", values_to = "prop_events") %>%
  left_join(next_month) %>%
  mutate(events = ifelse(total_events > rcd_limit, rcd_limit, total_events), events_nextmonth = ifelse(total_events_nextmonth > rcd_limit, rcd_limit, total_events_nextmonth))



fullOB_Rt_summary <- full_df %>%
  mutate(month_year = ym(paste0(year(date),"-",month(date)))) %>%
  group_by(month_year) %>%
  summarize(full_est = mean(est))
  

Rt_month_summary <- merged %>%
  filter(rcd_limit != 24) %>%
  mutate(month_year = ym(paste0(year(date),"-",month(date)))) %>%
  group_by(month_year, rcd_limit) %>%
  summarize(est = mean(est)) %>% #note this is taking the mean over dates and does not include weights for the number of people who contribute to Rt for a given time
  full_join(months_df) %>%
  left_join(fullOB_Rt_summary) %>%
  mutate(dRt = est - full_est, rel_dRt = dRt/full_est)

```
## Exploring some plots
```{r}
ggplot(months_df, aes(x = month_year, y = prop_events, color = as.numeric(rcd_limit))) +
  geom_point() +
  scale_color_viridis_c() +
  theme_bw()

ggplot(Rt_month_summary , aes(x = prop_events, y = dRt, color = as.numeric(rcd_limit))) +
  geom_point() +
  scale_color_viridis_c() +
  geom_hline(yintercept = 0) +
  theme_bw()
#no clear patterns

ggplot(Rt_month_summary , aes(x = prop_events_nextmonth, y = dRt, color = as.numeric(rcd_limit))) +
  geom_point() +
  scale_color_viridis_c() +
  geom_hline(yintercept = 0) +
  theme_bw()


ggplot(Rt_month_summary, aes(x = prop_events_nextmonth - prop_events, y = dRt, color = as.numeric(rcd_limit))) +
  geom_point() +
  scale_color_viridis_c() +
  geom_hline(yintercept = 0) +
  theme_bw()


ggplot(Rt_month_summary, aes(x = month_year, y = prop_events_nextmonth/prop_events, color = as.factor(as.numeric(rcd_limit)))) +
  geom_line() +
  scale_color_viridis_d() +
  theme_bw()



ggplot(NULL, aes(x,y))+
  geom_line(Rt_month_summary, mapping = aes(x = month_year, color = as.factor(as.numeric(rcd_limit)), y = prop_events_nextmonth-prop_events))+
  geom_point(merged %>% filter(rcd_limit != 24) %>% mutate(dist = (est - full_est)) %>% filter(!is.na(dist)), mapping = aes(x = date, y = dist/3, color = rcd_limit), alpha = 0.1) +
  scale_y_continuous(name = "prop_events_nextmonth-prop_events", sec.axis = sec_axis(trans=(~.*3), name = "dRt")) +
  scale_color_viridis_d() +
  theme_bw()+
  labs(x = "Date", y = "prop_nextmonth - prop_thismonth", color = "rcd limit")

ggplot(Rt_month_summary, aes(x = month_year, y = as.numeric(events_nextmonth)-as.numeric(events), color = as.factor(as.numeric(rcd_limit)))) +
  geom_line() +
  scale_color_viridis_d() +
  theme_bw()

```

# Trying to understand where the Turkana dataset falls
```{r}
turkana_facility_counts <- read.csv("data/turkana_facility_monthly_cases_cleaned.csv") %>%
  select(periodcode, organisationunitcode, total_positive) %>%
  mutate(month = ym(periodcode), assoc_facility = ifelse(organisationunitcode == 15661, "StMonica", ifelse(organisationunitcode == 18360, "Ngiitakito", ifelse(organisationunitcode == 14838, "Kerio", ifelse(organisationunitcode == 18354, "Nakechichok", ifelse(organisationunitcode == 15275, "Nadoto", ifelse(organisationunitcode == 15662, "StPatrick", NA)))))))


study_facility_counts <- nodes_pcrpos %>%
  filter(case_type == "facility") %>%
  group_by(assoc_facility, date) %>%
  count()

study_facility_counts <- data.frame(study_facility_counts) %>%
  rename("study_positive" = "n") %>%
  mutate(date = ymd(date)) %>%
  left_join(turkana_facility_counts %>% rename("date" = "month") %>% select(date, assoc_facility, total_positive)) %>%
  mutate(prop = study_positive/total_positive) %>%


study_counts <- study_facility_counts %>%
  group_by(date) %>%
  summarize(total_positive = sum(total_positive, na.rm = TRUE), study_positive = sum(study_positive, na.rm = TRUE)) %>%
  mutate(prop = study_positive/total_positive)

ggplot(study_facility_counts, aes(x = date, y = prop, color = assoc_facility)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5) +
  geom_line(study_counts, mapping = aes(x = date, y = prop), color = "black") +
  geom_point(study_counts, mapping = aes(x = date, y = prop), color = "black") +
  scale_color_locuszoom() +
  theme_bw()

sum(study_counts$study_positive, na.rm = TRUE)/sum(study_counts$total_positive, na.rm = TRUE)

```

